Community biomass and species richness temporal trends reveal contrasting

### 2021-12-13 paper/manuscript.Rmd
Methods reviewed

---
title: "Food-web temporal changes according to changes in species
richness and community biomass reveal contrasting assembly processes"
author: Alain Danet, Loubna El Madouri, Maud Mouchet, Elisa Thébault and Colin Fontaine (ordre non fixé)
date: \today 
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
  bookdown::word_document2:
fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \usepackage{float}
   - \usepackage[table]{xcolor}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
number_sections: true
---

@@

@@

Multiple studies reported dramatic declines of population and community biomass
[@ceballos_biological_2017; @hallmann_more_2017; @albano_native_2021], as well
as decrease in species richness [@barnosky_has_2011; @newbold_global_2015] and
turnover in species composition [@comte_climate_2021]. The documentation of
biodiversity temporal trends, is crucial to evaluate the effect of
global changes [but see @cardinale_is_2018]. Unfortunately, studies often
report trends of one dimension of biodiversity at a time [@hallmann_more_2017,
etc, but see @]. Because the functioning of ecosystem depends ultimately on
multiple facets of biodiversity, such as species richness, evenness and
functional composition [REF @thibaut_understanding_2013], the monitoring of one
biodiversity trend is not enough to project potential effects on ecosystem
functioning.

## Predictions come mainly from space-for-time substitution

Another concern is that the links between the different components of community
structure often come from spatial comparison [@dunne_network_2006;
@soininen_relationship_2012; @grace_integrative_2016 @danet_species_nodate]. In
order to use spatial comparison to predict the fate of communities facing
changes in species richness or community biomass, studies often implicitly use
the "space for time substitution" hypothesis. This method began to be widely
used in the early 20th century to infer the temporal succession of vegetation
from space variations [@pickett_space-for-time_1989]. It supposes a homology
between variations in space and in time i.e. that spatial sequences correspond
to temporal sequences.  This method have raised criticisms
[@damgaard_critique_2019], with studies showing analogous patterns in space and
time [@blois_space_2013] or not [@johnson_testing_2008;
@adler_contrasting_2007]. Spatial turnover of species composition, dispersal
limitation, disturbance and stochastic processes have been hypothesised to be
responsible for the failure of space for time substitution
[@adler_contrasting_2007; @blois_space_2013].

@@

The documentation of biodiversity trends generally neglect species interactions
although they are fundamentals to ecosystem functioning
[@thebault_food-web_2003; oconnor_general_2016]. In food-webs, empirical and
theoretical studies reported generally that trophic level positively scales to
community biomass [@wang_biodiversity_2018; @danet_species_nodate] (add
mechanisms ?). Trophic level has also been shown to stabilise the temporal
variation of community biomass [@shanafelt_stability_2018;
@danet_species_nodate]. Connectance, which measures the degree of generalism in
a network, was shown to increase robustness to extinction [@dunne_network_2002].
However, the temporal variation of food-web structure is largely unknown
[@kortsch_disentangling_2021]. Food-web temporal variation are often reported
across small time-steps [@warren_spatial_1989; @pecuchet_novel_2020;
@kaartinen_high_2012].

We know very little how food-web structure vary across time and particularly how
those variations are linked to the variations of other aspects of community
structure such as species richness and community biomass. Spatially, connectance
has been shown to be negatively related to species richness in empirical
food-webs [reviewed in @dunne_network_2006]. Connectance has been shown to scale
positively or negatively with community biomass in theoretical models
[@thebault_food-web_2003; @poisot_trophic_2013] (*Add mechanism?*). In order to
improve our understanding on the assembly mechanisms in food-webs, it is thus
crucial to assess how biomass and species richness variation are associated to
variations in food-web structure.

In this line, how the temporal changes of community biomass are related to
temporal changes of species richness can still be question in natural settings,
especially in food-webs. Our knowledge on how the temporal changes of food-web
structure are related  to temporal changes in community biomass and species
richness is almost zero. Our knowledge on the relationships between food-web
structure comes often from comparison of food-web snapshots along spatial
gradient, raising the question of the validity of space for time substitution to
infer consequences of species richness and community biomass on food-web
structure. The increasing amount of availability ecological temporal series
makes possible to assess those questions. Using a standardized monitoring of
fish freshwater communities on the period 1995-2018 and a method of food-web
inference based on allometric relationships and food diet data, we assessed the
relationships among temporal changes of food-web structure, trends of community
biomass and trends in species richness. We then assessed the validity of space
for time substitution by comparing if temporal relationships display the same
patterns as spatial gradient.

@@

@@

We selected the stations that have been monitored at least 10 times on the
1995-2018 period.

@@

We first inferred the body mass of each fish individual thanks to an allometric
law, $B_i = 0.01 \times l_i^3.03$, $B_i$ being biomass in gram and $l_i$ the
body length in millimetre of the individual $i$ [@froese_cube_2006]. Then the
biomass of each species and community biomass were obtained by summing the body
mass of individuals. Finally, the biomass was reported to the sampled surface
(in $m^2$) to account for sampling effort.

@@

We inferred food-web structure in two steps [@bonnaffe_comparison_2021;
@danet_species_nodate]: (1) we first computed the metaweb, describing the
potential trophic interactions between all species size classes, (2) then we
extracted each local time snapshot of food-webs by extracting species size
classes and their interactions according their presence in a given sampling
event.

We inferred food-web structure according to ontogenic diet data and body size
[@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019;
@bonnaffe_comparison_2021; @danet_species_nodate]. The method is extensively
described in @bonnaffe_comparison_2021 and @danet_species_nodate. The food diet
of stream fishes, as many organisms, change along their life and therefore
according to their body size. Each fish species was divided in nine size
classes, named trophic species, equally distributed in their observed body size
range.  A trophic species is a group of individuals of the same species and size
class that are expected to share the same preys and predators. We also added
seven nodes present in the ontogenic food diet database: detritus, biofilm,
phytoplankton, zooplankton, macrophages, phytobenthos and zoobenthos.  The
resulting metaweb contained 45 fish species divided into nine size classes plus
7 resource nodes, summing up to 412 nodes.  Consideration of size classes, i.e.
trophic species instead of species, improves inference of trophic interactions
[@jennings_weak_2001].

We determined the feeding interactions of trophic species according to their
species identity and their body size. The diet database described two or
three life stages by fish species, each one being delimited by body size
[@bonnaffe_comparison_2021; @danet_species_nodate]. The life stage of a fish
trophic species was determined by the centre of its body size range, hereafter
midpoint. The ontogenic fish diet database was filled according to fishbase
(Froese et al. 2021) and literature [@bonnaffe_comparison_2021;
@danet_species_nodate].

The feeding interactions between fishes were based on the body size ratio
between predators and preys. As this, we defined the predation window of a
piscivorous trophic species as 3% to 45% of its midpoint
[@mittelbach_ontogeny_1998; @claessen_impact_2002; @hart_handbook_2008].  A
trophic link was defined between the piscivorous trophic species and all the
trophic species whose midpoint in body size range was included in the predation
window.  The fish-resource trophic links were set between the trophic species
and the resources nodes according to the food items of their size class.
Finally, the feed interactions between resources were defined according to the
literature [@allan_stream_2007; @hart_handbook_2008].

## Food-web and community structure

We characterized food-webs by their connectance, describing the degree of
generalism in a network [@dunne_network_2006; @delmas_analysing_2019]. It was 
computed as the ratio between the actual number of links and the number of
potential links ($C = \frac{L}{S(S-1)}$, with $L$ being the number of links and
$S$ the number of trophic species). We also computed the average trophic level,
weighted by the biomass of the trophic species. The trophic level of a trophic
species was computed recursively as 1 added to the mean trophic level of the
preys that it consumes.

We described community structure by its community biomass, species richness and
the evenness of the number of individual and the amount of biomass across
species. The evenness was quantified with the Pielou index, computed as the
Shannon diversity index divided by the number of species ($P = \frac{H'}{N}$,
with $H'$ being the Shannon diversity index and $N$ the number of species).

@@

## Classification of temporal trajectories

To characterize the links among the temporal trends of network and community
structure and the temporal trends of community biomass and species richness, we
characterized the temporal trajectories of network and community structure at
each site. We used a method based on second order polynomial which allows to
assess both the direction and the monotony of temporal trends
[@rigal_method_2020]. To avoid the inclusion of many types of biomass and species
richness trends (i.e. accelerating, decelerating, etc...), we selected the
stations that displayed monotonic trends for these variables, i.e. the ones for
which the second order polynomial term was non significant. Over the `r nb_st`
stations, the selection procedure resulted in `r nb_st_rich_bm` for community
biomass and species richness analysis. However, the results were robust to the
inclusion of non monotonic trends (Fig. SXX).

There was a clear relationship between the size of the communities and their
temporal dynamics. Communities that displayed low community biomass at the
beginning of their monitoring tend to display biomass increase through time
and reversely. We observed the same pattern for species richness, with
communities that displayed high species richness at the beginning of their
monitoring tend to display a decrease in species richness through time and
reversely.

@@

In spatial analysis, we selected the same stations that were used for the
temporal analysis. We summarised temporal series of each stations by the median
value over time of each variable.

@@

## Temporal trends

We modelled the dynamics the temporal trends of food-web structure ($dN$)
in function of the temporal trends of community biomass ($dB$) and species
richness ($dR$) in order to test the relationship among the temporal dynamics of network
structure from one side and species richness and community biomass from another
side.

$dN = \alpha + \beta_1dB + \beta_2dR + \epsilon$

@@

  - $dN$: temporal trend of network structure and communities structure
  - $dB$: temporal trend of community (log) biomass
  - $dR$: temporal trend of (log) species richness

In order to control for the uncertainties of the estimated temporal trends, each
temporal trend was weighted by the inverse of the product of the standard errors of the
estimated trends, i.e. slopes for the response and the explicative
variables. We also tested a statistical model taking in account the interaction
between the temporal trends of species richness and community biomass (i.e.
adding $\beta_3dBdR$), the results are presented in supplementary material (Fig.
SXX). To model the relationships between the temporal trends of community
biomass and the one of species richness, we built equations as $dB = \alpha +
\beta_1dR + \epsilon$.

## Spatial trends

Spatial gradient of food-web structure are partly driven by environmental gradients, so
we controlled relationships among community structure variables by two PCA axis
synthesizing respectively (1) stream size and (2) average temperature and
enrichment. Enrichment was characterized by the average Biological Oxygen
Demand [BOD, @mallin_factors_2006].

$N = \alpha + \beta_1B + \beta_2R + \beta_3E_1 + \beta_4E_2 + \epsilon$

@@

  - $N$: median network structure and communities structure
  - $B$: median community biomass
  - $R$: median species richness 
  - $E_1$: PCA1 related to stream size
  - $E_2$: PCA2 related to average temperature and average BOD

To model the relationships between community biomass and species richness along
the spatial gradient, we built the equation as $B = \alpha + \beta_1R +
\beta_3E_1 + \beta_4E_2 + \epsilon$.

The significativity of the effects were tested with ANOVA. In order to compare
the magnitude of the effects of community biomass and species richness in both
temporal and spatial models, we computed standardized slopes ($r_{\delta}$) and
their associated standard errors.

$r_{\delta} = \beta\frac{s_x}{s_y}$

With $\beta$ being the unstandardised slope, $s_x$ and $s_y$ the standard
deviation of the explicative and response variable.

@@

sum(stable_st_var[["log_bm_std"]] %in% stable_st_var[["log_rich_std"]] == TRUE) / length(stable_st_var[["log_bm_std"]])

sum(stable_st_var[["log_rich_std"]] %in% stable_st_var[["log_bm_std"]] == TRUE) / length(stable_st_var[["log_rich_std"]])
```

@@

rsq_piel_bm_sp <- sp_rsq_table_bm_rich_mono_stable_trends %>%
  filter(Response == "piel_bm") %>%
  .$Conditional %>%
  round(., 3)
```

Overall, we found high heterogeneity in the explicative power of statistical
models linking temporal dynamics of food-web structure and the one of community
biomass and species richness spanning from $R_{adj}^2= `r rsq_piel_nind`$ for
the evenness of abundances to $R_{adj}^2= `r rsq_tlvl`$ for average trophic
level (Table \@ref(tab:rsq)). The spatial models were more performant, with
$R_{cond}^2$ varing from $`r rsq_tlvl_sp`$ to $`r rsq_piel_bm_sp`$ (Table
\@ref(tab:rsq-sp)). Over the station in the dataset, we found that
`r res_trends[res_trends$variable == "log_bm_std", ]$total[1]` stations displayed monotonic trends for biomass. Among them, 
`r res_trends[res_trends$variable == "log_bm_std" & res_trends$direction == "stable", ]$percent`%
were classified stable,
`r res_trends[res_trends$variable == "log_bm_std" & res_trends$direction == "decrease", ]$percent`%
decreasing,
`r res_trends[res_trends$variable == "log_bm_std" & res_trends$direction == "increase", ]$percent`%
increasing. Regarding the species richness, we found that
`r res_trends[res_trends$variable == "log_rich_std", ]$total[1]` stations displayed monotonic trends.
`r res_trends[res_trends$variable == "log_rich_std" & res_trends$direction == "stable", ]$percent`%
were classified stable,  
`r res_trends[res_trends$variable == "log_rich_std" & res_trends$direction == "decrease", ]$percent`%
as decreasing, and 
`r res_trends[res_trends$variable == "log_rich_std" & res_trends$direction == "increasing", ]$percent`% increasing.

---
#Overall, we found no significant effect of the temporal gain or losses of
#communinity biomass biomass and species richness on the relationship between the
#temporal trends of community structure and that of biomass and species richness
#(Table \@ref(tab:aov)).
---

## Relationships between temporal trends

@@

We found a strong positive relationship between the temporal trends of community
biomass and the one of species richness ($r_\delta = `r tps_r_sest["p_bm"]`$,
Fig. \@ref(fig:bef), Table \@ref(tab:aov)), meaning that both temporal trends
had the same direction in average. At stations
where species richness decreased over time, community biomass also decreased
over time. On the other hand, at stations where species richness increased over
time, community biomass also increased over time. In spatial analysis, we also
found a positive relationship between median community biomass and median
species richness, but twice less strong than the temporal one
($r_\delta = `r sp_b_sest["rich"]`$).

The temporal trends of connectance was positively linked with the one of
community biomass ($r_\delta = `r tps_bm_sest["ct"]`$), meaning also that both
temporal trends had the same direction in average. That is, the stations where
food-web connectance decreased over time had also community biomass that
decreased over time and reversely. It means that communities that lose biomass
over time, also loses generalist predators and reversely that communities
gaining biomass over time also gain generalist predators. Spatially, we found
also a positive relationship between median connectance and median species
richness ($r_\delta = `r sp_b_sest["ct"]`$). However, we did not observe
significant relationships between the temporal trends of connectance and the one
of species richness (Fig. \@ref(fig:bmr)D) when we found spatially a stong
negative relationship between median connectance and median species richness
($r_\delta = `r sp_r_sest["ct"]`$, Fig. \@ref(fig:std)).

We found that the temporal trends of average trophic level had a strong positive
relationship with the one of community biomass 
($r_\delta = `r tps_bm_sest["tlvl"]`$, Fig. \@ref(fig:bmr)B, Table
\@ref(tab:aov)), meaning
that stations where community biomass decreased over time, average trophic level
decreased over time and vice versa. It means that in communities losing biomass
over time, the average biomass of the food-web moved toward lower trophic level
and vice versa. The spatial gradient also showed a positive relationship between
median average trophic level and median community biomass, but twice less strong
than the temporal one ($r_\delta = `r sp_b_sest["tlvl"]`$, Fig. \@ref(fig:std),
Table \@ref(tab:aov-sp)). As for connectance, the temporal trends of average
trophic level was unrelated to the one of species richness. It means in
communities either losing or gaining species richness over time, the location of
average biomass along   trophic level did not significantly change.  However, we
found a strong negative relationship along the spatial gradient 
($r_\delta = `r sp_r_sest["tlvl"]`$, Fig. \@ref(fig:std), Table
\@ref(tab:aov-sp)).

>>>>

---
title: "Community biomass and species richness temporal trends reveal contrasting
    assembly and disassembly processes in natural food-webs"
author: Alain Danet, Loubna El Madouri, Maud Mouchet, Elisa Thébault and Colin Fontaine (ordre non fixé)
date: \today 
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
  bookdown::word_document2:
fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \usepackage{float}
   - \usepackage[table]{xcolor}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
number_sections: true
---

@@

- Tentative titles:
  - How food-webs assembly and disassembly through time in natural settings?
    Lessons from freshwater fishes
  - Relationships among temporal trends of food-web structure, species richness
    and community biomass reveal contrasting assembly
    and disassembly processes
  - Temporal trends in food web structure according to richness and biomass
    reveal contrasting assembly and disassembly processes

@@

Multiple studies reported dramatic declines of population and community biomass
[@ceballos_biological_2017; @hallmann_more_2017; @albano_native_2021], as well
as decrease in species richness [@barnosky_has_2011; @newbold_global_2015] and
turnover in species composition [@comte_climate_2021]. Unfortunately, studies often
focus on the temporal trend of one dimension of biodiversity [@hallmann_more_2017,
etc, but see @]. Because the functioning of ecosystem depends ultimately on
multiple facets of biodiversity, such as species richness, evenness and
functional composition^[1]  [REF @thibaut_understanding_2013], the documentation
of the trend of one biodiversity facet is likely insufficient to evaluate
precisely its consequences
on ecosystem functioning.

[^1]: Commentaire de Colin: *je pense qu'il faut lister les facet de biodiv qu'on
va etudier. sans forcement s'y limiter?* Ici je pense que la composition en role
trophique rentre dans "functional composition" sans rentrer dans le détail.

## Current knowledge mainly comes from space-for-time substitution

Current knowledge on the links among different facets of biodiversity mainly
comes from from spatial comparison [@dunne_network_2006;
@soininen_relationship_2012; @grace_integrative_2016 @danet_species_nodate]. space-for-time substitution began to be widely
used in the early 20th century to infer the temporal succession of vegetation
from space variations [@pickett_space-for-time_1989]. It supposes a homology
between variations in space and in time, i.e. that spatial sequences correspond
to temporal sequences.  This method have raised criticisms
[@damgaard_critique_2019], with studies showing analogous patterns in space and
time [@blois_space_2013] or not [@johnson_testing_2008;
@adler_contrasting_2007]. Spatial turnover of species composition, dispersal
limitation, disturbance and stochastic processes have been hypothesised to be
responsible for the non-compliance of space-for-time substitution
[@adler_contrasting_2007; @blois_space_2013].

@@

The documentation of biodiversity temporal trends generally neglect species
interactions although they are fundamentals to ecosystem functioning
[@thebault_food-web_2003; @oconnor_general_2016]. In food-webs, spatial
comparisons in empirical and theoretical studies generally reported that trophic
level positively scales to community biomass [@wang_biodiversity_2018;
@danet_species_nodate]. Trophic level has also been shown to stabilise the
temporal variation of community biomass [@shanafelt_stability_2018;
@danet_species_nodate]. Connectance, which measures the degree of generalism in
a network, was shown to increase robustness to extinction [@dunne_network_2002].
However, the temporal variation of food-web structure is largely unknown
[@kortsch_disentangling_2021]. Food-web temporal variation are often reported
across small time-steps [@warren_spatial_1989; @pecuchet_novel_2020;
@kaartinen_high_2012]. Our knowledge are even more scarce on the variability of
food-web structure in time and particularly how those temporal variations are linked to
the temporal variations of other aspects of biodiversity such as species richness
and community biomass. Spatially, connectance has been shown to be negatively
related to species richness in empirical food-webs [reviewed in
@dunne_network_2006]. Connectance has been shown to scale positively or
negatively with community biomass in theoretical models
[@thebault_food-web_2003; @poisot_trophic_2013] (*Add mechanism?*). In order to
improve our understanding on the assembly mechanisms in food-webs, it is thus
crucial to assess how biomass and species richness temporal trends are associated to
temporal trends in food-web structure.

In this line, how the temporal changes of community biomass are related to
temporal changes of species richness is still questionned in natural settings,
especially in food-webs. Our knowledge on the relationships between food-web
structure and other facets of biodiversity often comes from the comparison of
food-web snapshots along spatial gradient, raising the question of the validity
of space-for-time substitution to infer consequences of species richness and
community biomass variations on food-web structure.

The increasing availability ecological temporal series and the development of
methods to reconstruct food-webs [@bonnaffe_comparison_2021;
@gravel_inferring_2013; @danet_species_nodate] makes possible to assess those
questions. We used the standardized monitoring of fish freshwater communities on
the period 1995-2018 over France, documenting the individual size of fishes in
400 sites during ten years or more. We then reconstruct around 4000 food-webs
based on allometric relationships and food diet data. We questionned i) how
temporal trends of food-web structure are  linked to trends of community biomass
and trends in species richness? ii) Are the relationships observed in time the
same as those observed in space, i.e. is the space-for-time substitution valid?

@@

We used a monitoring of freshwater fishes conducted all over France and
conducted by the Office Français de la Biodiversité (OFB, previously known as
ONEMA and AFB). The sampling protocols were standardized since 1995 and a
complete description of the sampling process can be found in
@danet_species_nodate (main text and appendixS2). The fishes are sampled using
electrofishing all across the streams in shallow streams and over the bank by
boat in deeper streams. All the fishes are counted and the bodysize of fishes are
recorded for a sample of them, enabling the estimate the distribution of
individual bodysize by species for each sampling [@bonnaffe_comparison_2021;
@danet_species_nodate].

@@

We selected the sites that have been monitored at least 10 years on the
1995-2018 period, with the same sampling protocol over time. We selected
one sampling a year by site. In each site, we ensured low temporal heterogeneity
in sampling period by selecting sampling that were performed in the same period,
leading to a mean standard deviation of sampled month in each site of three
weeks [@danet_species_nodate]. Those steps leads to the selection of 
`r nb_st` sites (Figure \@ref{fig:matmet}A).

@@

We first inferred the body mass of each fish individual considering an allometric
law, $B_i = 0.01 \times l_i^3.03$, $B_i$ being the biomass expressed in gram and $l_i$ the
body length of the individual $i$ expressed in millimetres  [@froese_cube_2006]. Then the
biomass of each species and community biomass were obtained by summing the body
mass of individuals. Finally, the biomass was reported to the sampled surface
(in $m^2$) to account for sampling effort.

@@

We inferred food-web structure in two steps [@bonnaffe_comparison_2021;
@danet_species_nodate]: (1) we first computed the metaweb, describing the
potential trophic interactions among all size classes species, (2) then we
extracted each yearly snapshot of food-webs by extracting species size
classes and their interactions according to their presence in a given sampling
event.

We inferred the metaweb according to ontogenic diet data and body size
[@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019;
@bonnaffe_comparison_2021; @danet_species_nodate]. The method is extensively
described in @bonnaffe_comparison_2021 and @danet_species_nodate. The food diet
of stream fishes changes along their life cycle or stage, therefore
according to their body size. Each fish species was divided in nine size
classes, named "trophic species", equally distributed in their observed body size
range. Thus, a trophic species is a group of individuals of the same species and size
class that are expected to share the same preys and predators. We also added
seven nodes present in the ontogenic food diet database: detritus, biofilm,
phytoplankton, zooplankton, macrophages, phytobenthos and zoobenthos.  The
resulting metaweb contained 45 fish species divided into nine size classes plus
7 resource nodes, summing up to 412 nodes.  Consideration of size classes, i.e.
trophic species instead of species, improves inference of trophic interactions
[@jennings_weak_2001].

We determined the feeding interactions of trophic species according to their
species identity and their body size. The diet database described two or three
life stages by fish species, each one being delimited by body size
[@bonnaffe_comparison_2021; @danet_species_nodate]. The ontogenic fish diet
database was filled according to fishbase (Froese et al. 2021) and literature
[@bonnaffe_comparison_2021; @danet_species_nodate].

The feeding interactions among fishes were based on the body size ratio
between predators and preys. We defined the predation window of a
piscivorous trophic species as 3% to 45% of its midpoint
[@mittelbach_ontogeny_1998; @claessen_impact_2002; @hart_handbook_2008].  A
trophic link was defined between the piscivorous trophic species and all the
trophic species whose midpoint in body size range was included in the predation
window.  The fish-resource trophic links were set between the trophic species
and the resources nodes according to the food items of their size class.
Finally, the feeding interactions between resources were defined according to the
literature [@allan_stream_2007; @hart_handbook_2008].

## Community and food-web structure

At each site, yearly food-webs were described by their community biomass,
species richness and the evenness of biomass across species. The evenness was
quantified with the Pielou index, computed as the Shannon diversity index
divided by the number of species ($P = \frac{H'}{N}$, with $H'$ being the
Shannon diversity index and $N$ the number of species). Yearly food-web
structure was described by their connectance, describing the degree of
generalism in a network [@dunne_network_2006; @delmas_analysing_2019]. It was
computed as the ratio between the number of actual links and the number of
potential links ($C = \frac{L}{S(S-1)}$, with $L$ being the number of links and
$S$ the number of trophic species). We also computed the average trophic level,
weighted by the biomass of the trophic species. The trophic level of a trophic
species was computed recursively as 1 added to the mean trophic level of the
preys that it consumes [@herendeen_ecological_2008].

@@

## Temporal trends of food-web

To characterize the links among the temporal trends of network and community
structure and the temporal trends of community biomass and species richness, we
characterized the temporal trajectories of food-web structure at each site. We
modeled a linear relationship between food-web structure and
time with a second order polynome, which allows to assess both the direction and the monotony of temporal
trends [@rigal_method_2020]. We selected the sites that displayed monotonic
trends for these variables, i.e. the ones for which the second order polynomial
term was non significant according to t-tests, to avoid the inclusion of many
types of biomass and species richness trends (i.e. accelerating, decelerating,
etc...), . Over the `r nb_st` monitored sites, the selection procedure resulted in `r
nb_st_rich_bm` sites for community biomass and species richness analysis. However, the
results were robust to the inclusion of non monotonic trends (Fig. SXX).

There was a clear relationship between the size of the communities and their
temporal dynamics. Communities that displayed low community biomass at the
beginning of their monitoring tend to display a biomass increase through time
and reversely, we observed the same pattern for species richness. We
log-transformed (natural logarithm) community biomass and species richness to model their trends,
which decreased the former effect (Fig. SXX). Temporal
trends of community biomass and species richness were thus expressed in rate of
change by year.

@@

To assess the validity of the space-for-time substitution hypothesis, we
compared the relationships among food-web structure in time to the ones observed
along spatial gradient. To do so, we selected the same sites that were used
for the temporal analysis. We summarised temporal series of each station by the
median value over time of each facet of food-web structure described above.

@@

## Relationships among temporal trends

We modelled the relationships among the temporal trends of food-web structure ($dN$)
depending on the temporal trends of community biomass ($dB$) and species
richness ($dR$) with a linear model. in order to test the relationship among the temporal of food-web
structure, that is connectance, average trophic level and evenness of biomass
from one side and species richness and community biomass from the other side as
follows:

$dN_j = \alpha + \beta_1dB_j + \beta_2dR_j + \epsilon_j$

@@

  - $dN$: temporal trends of food-web structure: connectance, average trophic
    level and evenness of biomass across species
  - $dB$: temporal trend of ($\log_e$) community biomass
  - $dR$: temporal trend of ($\log_e$) species richness
  - $j$: the site
  - $\alpha$: intercept 
  - $\beta$: slope 
  - $\epsilon$: gaussian error

In order to control for the uncertainties of the estimated temporal trends, each
temporal trend was weighted by the inverse of the product of the standard errors of the
estimated trends, i.e. slopes for the response and the explicative
variables ($w_j = \dfrac{1}{s_{r_j} s_{e,j}}$, $w_j$ being the weight, $s_{r_j}$ and $s_{e_j}$ being
respectively the standard errors on the temporal trends of the response and
explicative variables). We also included a statistical model including the
interaction between the temporal trends of species richness and community
biomass (i.e.
adding $\beta_3dB_jdR_j$), the results are presented in supplementary material (Fig.
SXX). To model the relationships between the temporal trends of community
biomass and the one of species richness, we built the model as $dB_j = \alpha_j +
\beta_1dR_j + \epsilon_j$.

## Relationships along spatial gradient

Spatial gradients of food-web structure are partly driven by environmental gradients, so
we controlled relationships among variables by adding two PCA axis
synthesizing respectively (1) stream size and (2) average temperature and
enrichment. Enrichment was characterized by the average Biological Oxygen
Demand [BOD, @mallin_factors_2006][^2].

[^2]: commentaire de colin: il faudrait introduire spatial gradient des l'intro
et cette partie est plutot sur les relationships among spatial gradient of biodiv facets

$N_j = \alpha + \beta_1B_j + \beta_2R_j + \beta_3E_{1,j} + \beta_4E_{2,j} + \epsilon_j + \epsilon_{\alpha|i}$

@@

  - $N_j$: median network structure and communities structure
  - $B_j$: median community biomass
  - $R_j$: median species richness 
  - $E_{1,j}$: PCA1 related to stream size
  - $E_{2,j}$: PCA2 related to average temperature and average BOD
  - $\epsilon_{\alpha|i}$: random effect of the hydrographic basin
    $i$ on $\alpha$

To model the relationships between community biomass and species richness along
the spatial gradient, we built the equation as $B_j = \alpha + \beta_1R_j +
\beta_3E_{1,j} + \beta_4E_{2,j} + \epsilon_j$.

The significativity of the effects were tested with ANOVA. In order to compare
the magnitude of the effects of community biomass and species richness in both
temporal and spatial models, we computed standardized slopes ($r_{\delta}$) and
their associated standard errors as $r_{\delta} = \beta\frac{s_x}{s_y}$.
With $\beta$ being the unstandardised slope, $s_x$ and $s_y$ the standard
deviation of the explicative and response variables.

@@

#sum(stable_st_var[["log_bm_std"]] %in% stable_st_var[["log_rich_std"]] == TRUE) / length(stable_st_var[["log_bm_std"]])

#sum(stable_st_var[["log_rich_std"]] %in% stable_st_var[["log_bm_std"]] == TRUE) / length(stable_st_var[["log_rich_std"]])
```

@@

rsq_piel_bm_sp <- sp_rsq_table_bm_rich_mono_stable_trends %>%
  filter(Response == "piel_bm") %>%
  .$Conditional %>%
  round(., 3)
```
Over the sampling sites in the dataset, we found that
`r res_trends[res_trends$variable == "log_bm_std", ]$total[1]` stations displayed monotonic trends for biomass. Among them, 
`r res_trends[res_trends$variable == "log_bm_std" & res_trends$direction == "stable", ]$percent`%
were classified as stable,
`r res_trends[res_trends$variable == "log_bm_std" & res_trends$direction == "decrease", ]$percent`%
as decreasing,
`r res_trends[res_trends$variable == "log_bm_std" & res_trends$direction == "increase", ]$percent`%
as increasing. Regarding the species richness, we found that
`r res_trends[res_trends$variable == "log_rich_std", ]$total[1]` stations displayed monotonic trends.
`r res_trends[res_trends$variable == "log_rich_std" & res_trends$direction == "stable", ]$percent`%
were classified as stable,  
`r res_trends[res_trends$variable == "log_rich_std" & res_trends$direction == "decrease", ]$percent`%
as decreasing, and 
`r res_trends[res_trends$variable == "log_rich_std" & res_trends$direction == "increase", ]$percent`% as increasing.

Overall, we found high heterogeneity in the explicative power of statistical
models linking temporal dynamics of food-web structure, evenness of biomass  and the one of community
biomass and species richness spanning from $R_{adj}^2= `r rsq_piel_nind`$ for
the evenness of abundances to $R_{adj}^2= `r rsq_tlvl`$ for average trophic
level (Table \@ref(tab:rsq)). The models over the spatial gradient were more performant, with
$R_{cond}^2$ varing from $`r rsq_tlvl_sp`$ to $`r rsq_piel_bm_sp`$ (Table
\@ref(tab:rsq-sp)), $R_{cond}^2$ being the $R^2$conditional to the random effect. 
---
#Overall, we found no significant effect of the temporal gain or losses of
#communinity biomass biomass and species richness on the relationship between the
#temporal trends of community structure and that of biomass and species richness
#(Table \@ref(tab:aov)).
---

@@

We found a strong positive relationship between the temporal trends of community
biomass and the ones of species richness ($r_\delta = `r tps_r_sest["p_bm"]`$,
Fig. \@ref(fig:bef), Table \@ref(tab:aov)), meaning that both temporal trends
had the same direction in average. At the sites 
where species richness decreased over time, community biomass also decreased
over time. On the other hand, at the sites where species richness increased over
time, community biomass also increased over time. Along the spatial gradient, we also
found a positive relationship between median community biomass and median
species richness, but half the strength as the temporal one
($r_\delta = `r sp_b_sest["rich"]`$).

The temporal trends of connectance were positively linked with the ones of
community biomass ($r_\delta = `r tps_bm_sest["ct"]`$), also meaning that both
temporal trends had the same direction in average. That is, the sites where
food-web connectance decreased over time also had community biomass that
decreased over time and reversely. It suggests that communities that loose biomass
over time, also loose generalist predators and reversely that communities
gaining biomass over time also gain generalist predators. Spatially, we also found
a positive relationship between median connectance and median species
richness ($r_\delta = `r sp_b_sest["ct"]`$). However, we did not observe
significant relationships between the temporal trends of connectance and the
ones of species richness (Fig. \@ref(fig:bmr)D) while we found a stong
negative relationship between median connectance and median species richness
along the spatial gradient
($r_\delta = `r sp_r_sest["ct"]`$, Fig. \@ref(fig:std)).

We found that the temporal trends of average trophic level had a strong positive
relationship with the ones of community biomass 
($r_\delta = `r tps_bm_sest["tlvl"]`$, Fig. \@ref(fig:bmr)B, Table
\@ref(tab:aov)), meaning
that sites where community biomass decreased over time, average trophic level
decreased over time and *vice versa*. It suggests that in communities loosing biomass
over time, the average biomass of the food-web moved down to lower trophic level
and *vice versa*. The spatial gradient also showed a positive relationship between
median average trophic level and median community biomass, but half the strength 
as the temporal one ($r_\delta = `r sp_b_sest["tlvl"]`$, Fig. \@ref(fig:std),
Table \@ref(tab:aov-sp)). The temporal trends of average
trophic level was unrelated to the ones of species richness. It suggests that in
communities either loosing or gaining species richness over time, the location of
average biomass along trophic level did not significantly change according to species richnnes.  However, we
found a strong negative relationship in space 
($r_\delta = `r sp_r_sest["tlvl"]`$, Fig. \@ref(fig:std), Table
\@ref(tab:aov-sp)).

Evenness of biomass was negatively linked to community
biomass in time and in space 
($r_\delta = `r tps_bm_sest["p_bm"]`$ and $`r sp_b_sest["p_bm"]`$ resp.; Fig.
\@ref(fig:bmr)C, Fig. \@ref(fig:std)). Regarding time, it means that food-webs that
lost biomass over time also had a more even distribution of biomass across species over
time and *vice versa*. A similar reasoning applies in space. On the
other hand, we found a positive relationship between the temporal trends of
biomass evenness and the one of species richness (Fig. \@ref(fig:bmr)F, Table
\@ref(tab:aov), $r_\delta = `r tps_r_sest["p_bm"]`$). We found the same
relationship in space, twice as strong however 
($r_\delta = `r sp_r_sest["p_bm"]`$).
